### 2017-05-10 letter to CJM.tex
Wrapping up.

\documentclass[fontsize=10pt, paper=a4]{scrlttr2}

% Dont forget to read the KOMA-Script documentation, scrguien.pdf

\setkomavar{fromname}{Clifton Cunningham and David Roe} % your name
%\setkomavar{fromaddress}{}

\setkomavar{signature}{CC \& DR} % printed after the \closing
\renewcommand{\raggedsignature}{\raggedright} % make the signature ragged right

\setkomavar{subject}{Regarding CJM: MS 1903 (Commutative character sheaves and geometric types for supercuspidal representations)} % subject of the letter

\begin{document}
\begin{letter}{To: Canadian Journal of Mathematics}

\opening{Dear Isolde Prince,}  % eg. Hello

\closing{Best regards} %eg. Regards

\end{letter}
\end{document}