### 2020-10-20 referee_response.tex
Remove todo notes, add referee response

\documentclass[11pt]{article}
\usepackage{amsmath,amssymb,amsfonts,amsthm}

\newcommand{\kbar}{\overline{k}}
\newcommand{\A}{\mathbb{A}}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\Qp}{\Q_p}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\N}{\mathbb{N}_{\rm div}}
\newcommand{\Zp}{\Z_p}
\newcommand{\F}{\mathbb{F}}
\newcommand{\Fp}{\F_p}
\newcommand{\Fq}{\F_q}
\newcommand{\Fqbar}{\overline{\F}_q}
\newcommand{\Fqm}{\F_{q^m}}
\newcommand{\Fqn}{\F_{q^n}}
\renewcommand{\O}{\mathcal{O}}
\newcommand{\dom}{\operatorname{dom}}
\newcommand{\Spec}{\operatorname{Spec}}

\title{Response to Referee -- \\ Commutative character sheaves and geometric types for supercuspidal representations}
\author{Clifton Cunningham and David Roe}
\date{\today}

\begin{document}
\maketitle

We would like to thank the referee for a thorough and considered review.  The comments were very helpful and allowed us to improve the paper.

In addition to the line items suggested by the referee, we have made the following substantive changes.

\begin{enumerate}
\item We have added a reference to Fintzen's paper as suggested, and provided a more precise condition on $p$ (1,2).
\item We added precise references to the work of Lusztig that we depend on (11,20,22).
\item We replaced $p$-adic field with local field as suggested (15).
\item We agree that it would be nice if the genericity condition on $\phi_i$ could be translated to an illuminating condition on $\varphi_i$.  Unfortunately, while the condition can certainly be translated to the Greenberg transform side of the story, it doesn't seem to provide any insight.  We have therefore not addressed the referee's suggestion (16).
\item We have added a reference to a paper of Yu (18).
\item In addressing (19,20), we realized that hypothesis \textbf{H0} was unnecessary (it resulted from our misreading of one of Lusztig's results) so we have removed it.  We have added two paragraphs illustrating hypothesis \textbf{H1(b)}, giving examples both of representations that do not satisfy it and a broad class of representations that do discussed in a paper of Kaletha.
\item We added a reference to Hakim's paper in the introduction (24).
\end{enumerate}

\end{document}